Suitability of Graph Neural Network for Shape Memory Alloys

### 2024-05-24 src/tex/ms.tex
Tweaks

% Abstract with filler text
\begin{abstract}
	Shape Memory Alloys (SMAs) such as NiTi are pivotal in vibrational
	dampening, robotic arms, prosthetic hands, and mechanical vales as a
	result of their unique displacive phase transformation
	properties. However, assessing their phonon characteristics, which
	provide insight into the displacive phase transformation, by using
	conventional Density Functional Theory (DFT) methods is
	computationally challenging. In this paper we assess the use of
	recent Graph Neural Network (GNN) potentials as a viable
	alternative. These GNN potentials have leveraged large DFT-computed
	databases for training and provide a good approximation to the
	potential energy surfaces which enable inference on energies and
	forces in lattice and atomic dynamic calculations. We compare GNN
	potentials with the Embedded Atom Method (EAM) and Modified EAM
	style potentials in predicting phonon properties of NiTi and
	PtTi. We find that in general the EAM/MEAM potentials are slightly
	more performant than the GNN in that they appear to more accurately
	describe equation of state, phonon dispersions, and elastic
	constants. However, the MACE potential shows significant promise as
	an exploratory potential for designing and characterizing more
	complex SMAs. Our findings suggest GNN potentials as a promising
	initial avenue for advancing SMA research, in particular ternary and
	quaternary systems, upon fine-tuning of pre-trained GNN models with
	more relevant DFT data (i.e., force-constants) to permit better
	outcomes.
\end{abstract}

@@

\section*{Acknowledgements}

>>>>

% Abstract with filler text
\begin{abstract}
	Shape Memory Alloys (SMAs), such as NiTi, are crucial in vibration dampening, robotic arms, prosthetic hands, and mechanical valves due to their unique displacive phase transformation properties. However, assessing their phonon characteristics using conventional Density Functional Theory (DFT) methods is computationally expensive. This paper evaluates the use of recent Graph Neural Network (GNN) potentials as a viable alternative. These GNN potentials, trained on extensive DFT-computed databases, approximate potential energy surfaces, enabling inference on energies and forces in lattice and atomic dynamics calculations. We compare GNN potentials with the Embedded Atom Method (EAM) and Modified EAM (MEAM) potentials in predicting phonon properties of NiTi and PtTi. Generally, EAM/MEAM potentials appear to slightly outperform GNNs phonon, providing more accurate descriptions of the equation of state, phonon dispersions, and elastic constants. However, the MACE potential shows significant promise as an exploratory tool for designing and characterizing more complex SMAs. Our findings suggest GNN potentials as a promising initial avenue for advancing SMA research, particularly for ternary and quaternary systems, with improved outcomes possible by fine-tuning pre-trained GNN models with more relevant DFT data (e.g., force-constants).
\end{abstract}

@@

\section*{Acknowledgements}
The author would like to thank Prof.Venkateswara Manga at the University of Arizona for fruitful discussions.
